---
title: "digitalLDSorteR: Introduction to deconvolution of bulk RNA-Seq samples based on Deep Learning"
author:
- name: First Author
  affiliation: First Author's Affiliation
- name: Second Author
  affiliation: Second Author's Affiliation
  email: corresponding@author.com
package: digitalDLSorteR
output:
  BiocStyle::html_document
    # css: "./css/style.css"
    toc: true
    toc_float: true
abstract: |
  Description of your vignette
vignette: |
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignettePackage{digitalDLSorteR}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
options(width = 70)
require(BiocStyle)
```


# Introduction to `digitalDLSorteR`package

`digitalDLSorteR` is a package that implements a Deep Learning based method to 
enumerate and quantify the cell type composition of a bulk RNA-Seq sample. Our 
method makes use of a Deep Neural Network (DNN) model to adjust any cell type
composition defined from a single-cell RNA-Seq (scRNA-Seq) data. 

The method implement a pipeline that starts from scRNA-Seq data and, after a few 
steps, a Deep Neural Network (DNN) model is trained with pseudo-bulk RNA-Seq 
samples with known composition of cell types. The trained model is able to 
deconvolute any bulk RNA-Seq sample, determining the proportion of the 
different cell types present in it. 

The package can be used in two ways: for deconvoluting bulk RNA-Seq
samples using a pre-trained model provided by us or for building own models
trained from different scRNA-Seq samples with its own peculiarities. These 
new models can be published in order to make them available for other users
that work with similar data (e.g. neural environment, prostate cancer environment,
etc). For the moment, the available models allows the deconvolution of tumor
infiltrated lymphocytes (TILs) from breast cancer and colorectal cancer 
environments classified by our team.

# Instalation of package



# Use `digitalDLSorteR` for deconvolution of samples using pre-trained models

This is the simplest way to use the method: from a bulk RNA-Seq sample matrix
with genes as rows and samples as columns, perform the deconvolution of cell types
using a pre-trained model available in the package. As mentioned above, for the 
moment only two pre-trained models are available: breast cancer (`breast.chung`)
and colorectal cancer (`colon.li`) models. These two models are generated from 
two scRNA-Seq datasets that have been analyzed by our team:

* Breast cancer environment has been generated from GSE75688 (Chung et al., 2017). 
* Colorectal cancer environment has been generated from GSE81861 (Li et al., 2017).

These dataset have the particularity of presenting TILs and, therefore, the resulting
models are able to deconvolute immune cell types from bulk RNA-Seq samples. This 
is an advantage over other tools, since the model has been trained from scRNA-Seq
data from the tumor itself, preserving the specific characteristics of the tumor 
microenvironment as opposite to other approaches in which cells were isolated 
from blood. For knowing more about the datasets, see the provided documentation 
in the package. **Aquí realmente se puede comentar algo de las ventajas que 
presenta cada dataset y qué lo compone de forma resumida**.


## Example of the use of the package

Using the package in this way, it is only necessary to load the samples that you
want to deconvolute in R and use them as `data` argument for 
`deconvDigitalDLSorter` function. This object must be a `matrix` or a `data.frame`
with genes as rows and samples as columns. Genes must be annotated as SYMBOL, 
since the model was trained with this notation. Data must be TPMs, since model
was trained with this type of data. If data are not normalize, set `normalize`
argument equal to `TRUE` (by default).

Moreover, you must specify the model that will be used in `model` argument. 
In this example, we are going to deconvolute TCGA samples from colorectal cancer, 
so we have to use `colon.li` model setting `model = "colon"`:


```{r}
library(digitalDLSorteR)

deconvDigitalDLSorter(data = , model = "colon", batch.size = )
```




# Build your own model for deconvolution of bulk RNA-Seq samples in a determined 
environment


# Getting started

To enable the _Bioconductor_ style in your R Markdown vignette you need to include the following in the `DESCRIPTION` file:

    VignetteBuilder: knitr
    Suggests: BiocStyle, knitr, rmarkdown


# Style macros

_BiocStyle_ introduces the following macros for referring to _R_ packages:

* `r Biocpkg("IRanges")`, for _Bioconductor_ software, annotation and experiment data packages,
* `r CRANpkg("data.table")`, for _R_ packages available on CRAN,
* `r Githubpkg("rstudio/rmarkdown")`, for _R_ packages available on GitHub,
* `r Rpackage("MyPkg")`, for _R_ packages that are _not_ available on _Bioconductor_, CRAN or GitHub.


# Figures

Assign captions to figures in the code chunk option `fig.cap` to automatically number them, and to be able to reference them, see Figure \@ref(fig:plot). The figure label is generated from the code chunk label by prefixing it with `fig:`.

```{r plot, fig.cap="Regular figure. The first sentence of the figure caption is automatically emphasized to serve as figure title.", echo=FALSE}
plot(cars)
```

Small and wide figures can be specified by `fig.small` and `fig.wide` code chunk options.

```{r small, fig.cap="Small figure. A plot produced by a code chunk with option `fig.small = TRUE`.", fig.small=TRUE, echo=FALSE}
plot(cars)
```

```{r wide, fig.cap="Wide figure. A plot produced by a code chunk with option `fig.wide = TRUE`.", fig.wide=TRUE, echo=FALSE}
plot(cars)
```


# Equations

To number and reference equations, put them in equation environments and assign labels to them, see Equation \@ref(eq:binom).

\begin{equation}
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}
  (\#eq:binom)
\end{equation}


# Tables

Like figures, tables with captions will also be numbered and can be referenced, see Table \@ref(tab:table).

Fruit   | Price
------- | -----
bananas | 1.2
apples  | 1.0
oranges | 2.5

: (\#tab:table) A simple table. With caption.


# Cross-references

Apart from referencing figures (Section \@ref(figures)), tables (Section \@ref(tables)), and equations (Section \@ref(equations)), you can also use the same syntax to refer to sections by their default labels generated by pandoc.


# Side notes

Footnotes are displayed as side notes on the right margin^[this is a side note entered as a footnote], which has the advantage that they appear close to the place where they are defined.


# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
